### 2015-07-14 paper/Results.tex
Upload tables and figures

\documentclass[a4paper,12pt]{article}
\usepackage[german,english]{babel}
\usepackage{charter}
\usepackage[T1]{fontenc} 
\usepackage{amsmath,amssymb,amsthm} % Typical maths resource packages
\usepackage{amsfonts}
\usepackage{graphicx} % Packages to allow inclusion of graphics
\usepackage[authoryear]{natbib} % literature reference style
\usepackage{cancel}
\usepackage{dsfont}
\usepackage{amsthm}
\usepackage{color}
\usepackage{float}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{setspace}
\usepackage{rotating}
\usepackage[automark]{scrpage2}
\pagestyle{scrheadings}
\automark{section}
\clearscrheadings
%\ohead{\headmark} % puts chapter heading on top of each page
\bibliographystyle{ecta} % define citation style
% define page size, margin size
\setlength{\headheight}{1.8\baselineskip}
\voffset=-3cm
\hoffset=-3cm
\textheight24cm
\textwidth16.5cm
\topmargin1cm
\oddsidemargin3cm
\evensidemargin3cm
\renewcommand{\baselinestretch}{1.5} % define line line spacing = 1.5
\newcommand{\E}{\mathbb{E}}
\newcommand{\Lagr}{\mathcal{L}}
\usepackage{hyperref}

\begin{document}

\begin{center}
\textbf{Results Overview}
\end{center}

This document collates the results obtained in the file \texttt{gudat\_weldzius\_2014.R}. The following results are available:

\begin{itemize}
	\item Table \ref{cis_price_sr}: Chen et al. (2011), Prices Short Run
	\item Table \ref{cis_prod_sr}: Chen et al. (2001), Productivity Short Run
	\item Table \ref{cis_mark_sr}: Chen et al. (2011), Markups Short Run
	\item Table \ref{cis_price_lr}: Chen et al. (2011), Prices Long Run
	\item Table \ref{cis_prod_lr}: Chen et al. (2011), Productivity Long Run
	\item Table \ref{cis_mark_lr}: Chen et al. (2011), Markups Long Run
\end{itemize}

\begin{table}
\input{cis_price_sr.tex}
\caption{Prices, Short Run}
\label{cis_price_sr}
\end{table}

\begin{table}
\input{cis_prod_sr.tex}
\caption{Productivity, Short Run}
\label{cis_prod_sr}
\end{table}

\begin{table}
\input{cis_mark_sr.tex}
\caption{Markups, Short Run}
\label{cis_mark_sr}
\end{table}

\begin{table}
\input{cis_price_lr.tex}
\caption{Prices, Long Run}
\label{cis_price_lr}
\end{table}

\begin{table}
\input{cis_prod_lr.tex}
\caption{Productivity, Long Run}
\label{cis_prod_lr}
\end{table}

\begin{table}
\input{cis_mark_sr.tex}
\caption{Markups, Long Run}
\label{cis_mark_lr}
\end{table}

\begin{sidewaystable}
\input{sumstats.tex}
\caption{Summary Statistics}
\label{sumstats}
\end{sidewaystable}

\end{document}